Retrofitting Effect Handlers onto OCaml

### 2021-03-25 retro-concurrency.tex
british english to be consistent and expand acronym

Effect handlers~\cite{Plotkin09} provide a modular foundation for user-defined
effects. The key idea is to separate the definition of the effectful operations
from their interpretations, which are given by \emph{handlers} of the effects.
For example,
%
\begin{lstlisting}
effect In_line : in_channel -> string
\end{lstlisting}
%
declares an \emph{effect} |In_line|, which is parameterised with an input
channel of type |in_channel|, which when \emph{performed} returns a |string|
value. A computation can perform the |In_line| effect without knowing how the
|In_line| effect is implemented. This computation may be enclosed by different
handlers that handle |In_line| differently. For example, |In_line| may be
implemented by performing a blocking read on the input channel or performing
the read asynchronously by offloading it to an event loop such as |libuv|,
without changing the computation. Thanks to the separation of effectful
operations from their implementation, effect handlers enable new approaches to
modular programming. Effect handlers are a generalization of exception
handlers, where, in addition to the effect being handled, the handler is
provided with the delimited continuation~\cite{Danvy90} of the perform site,
which may be resumed later. This enables non-local control-flow mechanisms such
as resumable exceptions, lightweight threads, coroutines, generators and
asynchronous I/O to be composably expressed.

@@

We now present the implementation details of effect handlers in Multicore
OCaml. While we assume an x86\_64 architecture for the remainder of this paper,
our design does not preclude other architectures and OSs.

>>>>

Effect handlers~\cite{Plotkin09} provide a modular foundation for user-defined
effects. The key idea is to separate the definition of the effectful operations
from their interpretations, which are given by \emph{handlers} of the effects.
For example,
%
\begin{lstlisting}
effect In_line : in_channel -> string
\end{lstlisting}
%
declares an \emph{effect} |In_line|, which is parameterised with an input
channel of type |in_channel|, which when \emph{performed} returns a |string|
value. A computation can perform the |In_line| effect without knowing how the
|In_line| effect is implemented. This computation may be enclosed by different
handlers that handle |In_line| differently. For example, |In_line| may be
implemented by performing a blocking read on the input channel or performing
the read asynchronously by offloading it to an event loop such as |libuv|,
without changing the computation. Thanks to the separation of effectful
operations from their implementation, effect handlers enable new approaches to
modular programming. Effect handlers are a generalisation of exception
handlers, where, in addition to the effect being handled, the handler is
provided with the delimited continuation~\cite{Danvy90} of the perform site,
which may be resumed later. This enables non-local control-flow mechanisms such
as resumable exceptions, lightweight threads, coroutines, generators and
asynchronous I/O to be composably expressed.

@@

We now present the implementation details of effect handlers in Multicore
OCaml. While we assume an x86\_64 architecture for the remainder of this paper,
our design does not preclude other architectures and operating systems.
